#Reproducible Research, Project 1  
###Tobias Crabtree
###Coursera, August 2014
========================================================

##Loading and preprocessing the data in R
Personal movement activity data was downloaded August 2014 from 
https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip

The variables included in this dataset are:
- steps: Number of steps taking in a 5-minute interval (missing values are coded as NA)
- date: The date on which the measurement was taken in YYYY-MM-DD format
- interval: Identifier for the 5-minute interval in which measurement was taken

The dataset is stored in a comma-separated-value (CSV) file and there are a total of 17,568 observations in this dataset.

Transformations to the data include:
- Changed class to dates
- Removed missing data (i.e., steps has 2304 NA's)

```{r}
require(knitr)
data <- read.csv("activity.csv") #read data 
summary(data) # review data
data$date <- as.Date(data$date,"%Y-%m-%d") # transform dates
data <- na.omit(data) # remove missing data 

```

## What is the mean total number of steps taken per day?
```{r fig.width=7, fig.height=6}
require(plyr) # load required plyr package
dataSum <- ddply(data, .(date), summarize, total=sum(steps))
dev.copy(png,"totalsteps.png")
hist(dataSum$total, main="Total number of steps taken by day", 
     xlab="Total number of steps by day", col="white")
dev.off()

mean(dataSum$total)
median(dataSum$total)
```

## What is the average daily activity pattern?
```{r fig.width=7, fig.height=6}
require(plyr) # load required plyr package
dataMeanInt <- ddply(data, .(interval), summarize, mean=mean(steps)) # calculate mean steps by day
dev.copy(png,"avgdailyactivity.png")
plot(dataMeanInt$interval, dataMeanInt$mean, type="l", main="Average Daily Activity Pattern", 
     xlab="5 minute interval", ylab="Avg. # of Steps")
dev.off()
dataMeanInt2 <- ddply(data, .(interval, steps), summarize, mean=mean(steps)) # calculate mean steps by day
max <- max(dataMeanInt2$steps) # determine max steps
df <- dataMeanInt2[dataMeanInt2$steps==max,] # create dataframe with max steps and interval
interval <- df$interval
interval ## returns 5 minute interval with the max amount of steps
```


## Inputting missing values
```{r}
data2 <- read.csv("activity.csv") #re-read data as previously removed missing values
sum(is.na(data2)) # returns #'s of missing values

# fill in missing values with mean # of steps and create a new dataset
data2$steps[is.na(data2$steps)] <- mean(data2$steps, na.rm=TRUE)

# create a histogram
data2Sum <- ddply(data2, .(date), summarize, total=sum(steps))
dev.copy(png,"totalsteps2.png")
hist(data2Sum$total, main="Total number of steps taken by day", 
     xlab="Total number of steps by day", col="white")
dev.off()

mean(data2Sum$total)
median(data2Sum$total)

```
By inputting missing values the mean and median now converge to be the same #.

## Are there differences in activity patterns between weekdays and weekends?
```{r}
weekday <- subset(data, select=weekdays(date))
weekend <- 
```

